### 2022-02-26 covid_transparency_dataset_creation.Rmd

---
title: "Transparency indicators across the dentistry and oral health literature"
author: Eero Raittio, Ahmad Sofi-Mahmudi, Erfan Shamsoddin, Sergio E Uribe, Julie Farmer
date: "1/8/2022"
output: html_document
---

# Loading required packages

```{r}
pacman::p_load(tidyverse,
               rtransparent, 
               metareadr, 
               europepmc)
```

# Load the datasets

## Load all open access dentistry and oral health papers indexed in PubMed to the end of 2021:

```{r}
COVID_19_keywords <- read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vS6ibaTIJ0-4MiOZrLHKwo-Ya3MjlGX4f-dPQwdmWwQZhRGi6C83K1I_WceUJPH9zGyBc4L-lSalET_/pub?output=csv")
```

```{r}
COVID_19_keywords <- COVID_19_keywords %>% mutate(
  search.term.title = paste0("TITLE:", '"', Keyword, '"'),
  search.term.keyword = paste0("KW:", '"', Keyword, '"'),
  search.term.results = paste0("RESULTS:", '"', Keyword, '"'))

TitleQuery <- paste(COVID_19_keywords$search.term.title,
                     collapse = " OR ")

KeywordQuery <- paste(COVID_19_keywords$search.term.keyword,
                     collapse = " OR ")

ResultQuery <- paste(COVID_19_keywords$search.term.results,
                     collapse = " OR ")
```

## Load all open access dentistry and oral health papers indexed in PubMed to the end of 2021:

```{r}

db_2020_1 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2020-06-30]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")

db_2020_1 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2020-06-30]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")

db_2020_2 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-07-01 TO 2020-12-31]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")

db_2021_1 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2021-01-01 TO 2021-06-30]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")

db_2020_2 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2021-07-01 TO 2021-12-31]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")

db <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2021-12-31]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 1000000, output = "parsed")

db <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2021-12-31]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 1000000, output = "parsed")

db <- epmc_search(
  query = paste0(
    "'(",
    TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery,
    ") ",
    'AND (SRC:"MED") 
    AND (LANG:"eng" OR LANG:"en" OR LANG:"us") 
    AND (FIRST_PDATE:[2020-01-01 TO 2021-12-31])
    AND ((IN_EPMC:y) OR (OPEN_ACCESS:y))
    AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")',
    "'"
  ),
  limit = 100000, output = "parsed"
)
```

Let's see how many duplicates are there:
```{r}
table(duplicated(db$pmid))
```

Removing the duplicates:
```{r}
db <- db %>% distinct(pmid, .keep_all = TRUE)
```

Let's see the first six rows of the database:
```{r}
head(db)
```

Removing "PMC" from the cells
```{r}
db_2020_1$pmcid_ <- gsub("PMC", "", as.character(db_2020_1$pmcid))

db_2020_2$pmcid_ <- gsub("PMC", "", as.character(db_2020_2$pmcid))

db_2021_1$pmcid_ <- gsub("PMC", "", as.character(db_2021_1$pmcid))

db_2021_2$pmcid_ <- gsub("PMC", "", as.character(db_2021_2$pmcid))

db$pmcid_ <- gsub("PMC", "", as.character(db$pmcid))
```

Now, we make a folder for xml format articles and switch to that folder:
```{r}
dir.create("pmc")
setwd("pmc")
```

Next, we download xmls in format accessible with rtransparent:
```{r}
db_2020_1 <- subset(db_2020_1, pmcid_!=c(7264604, 7194244, 7323088, 7300787, 7276714, 7262104, 7228371))
db_2020_2 <- subset(db_2020_2, pmcid_!=c(7677857, 7164703, 7678408, 7486161, 7645956, 7486166, 7675701, 7405208, 7753328, 7575215, 7759253, 7568026, 7598225, 7416582, 7649313, 7492687, 7571788, 7675665, 7572235, 7457217))
db_2021_1 <- subset(db_2021_1, pmcid_!=c(8013335, 8013644, 8250125, 8014725, 8251275, 7486807, 8420411, 8251218, 8251305, 7894077))
db_2021_2 <- subset(db_2021_2, pmcid_!=8653067)

db <- subset(db_2021_2, pmcid_!=c(7264604, 8013335))

sapply(db_2020_1$pmcid_, mt_read_pmcoa)

sapply(db_2020_2$pmcid_, mt_read_pmcoa)

sapply(db_2021_1$pmcid_, mt_read_pmcoa)

sapply(db_2021_2$pmcid_, mt_read_pmcoa)

sapply(db$pmcid_, mt_read_pmcoa)

db <- subset(db, !(pmcid_ %in% c(8251275, 8251218, 7486161, 7228371, 8485571)))

```
Error in 8013644 - The metadata format 'pmc' is not supported by the item or by the repository.

Error in 7572235 - The value of the identifier argument is unknown or illegal in this repository.

Now we run rtransparent:
```{r}
filepath = dir(pattern=glob2rx("PMC*.xml"))

results_table_all <- sapply(filepath, rt_all_pmc)

results_table_data <- rt_data_code_pmc_list(
  filepath,
  remove_ns=F,
  specificity = "low")

write.csv(rt_data_code_pmc_list(filepath, remove_ns=F, specificity = "low"), "rt_data.csv")

f1_data <- rt_data_code_pmc_list(f1, remove_ns=F, specificity = "low")

```

A list is created now. We should convert this list to a dataframe:
```{r}
df <- data.table::rbindlist(results_table_all, fill = TRUE)
```

Merge data sharing results to database file:

```{r}
setwd('..')
db <- db[!duplicated(db[, 4]),]
opendata <- merge(db, results_table_data, by = "pmid") %>% merge(df)
View(opendata)
```