### 2022-04-27 rmarkdown-covid-19-transparent-analysis.Rmd

---
title: "Transparency of COVID-19-related research - Analysis"
author: Ahmad Sofi-Mahmudi, Eero Raittio, Sergio E. Uribe
date: "2022-02-11"
output: html_document
---

# Loading required packages

```{r}
pacman::p_load(dplyr,
               ggplot2,
               knitr,
               here,
               tidyr,
               europepmc,
               lubridate,
               ggpubr,
               epiR,
               gtsummary,
               forcats,
               nortest,
               caret,
               tibble)
```

# Loading the datasets

```{r}
opendata <- read.csv(here("data", "covid_transparency_opendata.csv"))
```

# Results
Number of all papers (open access and non open access), open-access papers and open-access percentage:
```{r}
kable(data.frame(hits_all = 175544, 
           hits_oa = 133467,
           oa_percentage = round((133467/175544)*100, 1)))
```

Undownloadable papers:
```{r}
c(number = 133467 - nrow(opendata), 
  percentage = round((133467 - nrow(opendata))/133467*100, 1))
```

First, adding the real publication year and month to the datasets. The real publication year/month is the year/month the paper was first appear online stored in firstPublicationDate column.
```{r}
opendata <- opendata %>%
        mutate(year_firstpub = year(as.POSIXlt(firstPublicationDate,
                                               format = "%Y-%m-%d")),
               month_firstpub = month(as.POSIXlt(firstPublicationDate,
                                                 format = "%Y-%m-%d")))
```

Number of open access papers per year:
```{r}
kable(table(opendata$year_firstpub))
```

The number of journals in our dataset:
```{r}
length(table(opendata$journalTitle))
```
Top 10 journals with the highest number of articles in our dataset, from high to low:
```{r}
kable(head(table(opendata$journalTitle) 
           %>% as.data.frame() 
           %>% arrange(desc(Freq)), 10))
```

Also, same table, but with all the journals

```{r}
opendata %>% 
        mutate(journal = fct_lump_prop(journalTitle, prop = 0.01) ) %>% 
        count(journal) %>% 
        arrange(desc(n)) %>% 
        knitr::kable()
```

The mean and the median of the number of citations to these references:
```{r}
kable(data.frame(Mean = round(mean(opendata$citedByCount),1),
           SD = round(sd(opendata$citedByCount), 2),
           Median = median(opendata$citedByCount),
           IQR = IQR(opendata$citedByCount)
           ))

```

Characteristics of the paper with the highest number of citations:
```{r}
kable(opendata[which.max(opendata$citedByCount),] %>% 
  select(citedByCount, pmid, pmcid, doi, title, authorString, journalTitle, year_firstpub))
```

The number and percentage of papers with a conflict of interest (CoI) disclosure:
```{r}
kable(data.frame(
  number = length(opendata$is_coi_pred[opendata$is_coi_pred == TRUE]),
           percentage = round(length(opendata$is_coi_pred[opendata$is_coi_pred == TRUE])/nrow(opendata)*100, 1)
           ))
```

Confidence interval for CoI:
```{r}

kable(round(epi.prev(pos = length(opendata$is_coi_pred[opendata$is_coi_pred == TRUE]),
         tested = nrow(opendata),
         se = 0.992,
         sp = 0.995)$ap, 
      1))
```

Comparing CoI disclosure rate in three years:
```{r}
kable(round(prop.table(table(opendata$is_coi_pred, opendata$year_firstpub), 2)*100, 1))
```

Chi-square test:
```{r}
chisq.test(opendata$is_coi_pred, opendata$year_firstpub)
```

The number and percentage of papers with a funding statement:
```{r}
kable(data.frame(number = length(opendata$is_fund_pred[opendata$is_fund_pred == TRUE]),
           percentage = round(length(opendata$is_fund_pred[opendata$is_fund_pred == TRUE])/nrow(opendata)*100, 1)
           ))
```

Confidence interval for funding statement:
```{r}

kable(round(epi.prev(pos = length(opendata$is_fund_pred[opendata$is_fund_pred == TRUE]),
         tested = nrow(opendata),
         se = 0.997,
         sp = 0.981)$ap, 
      1))
```

Comparing funding statement rate in three years:
```{r}
kable(round(prop.table(table(opendata$is_fund_pred, opendata$year_firstpub), 2)*100, 1))
```

Chi-square test:
```{r}
chisq.test(opendata$is_fund_pred, opendata$year_firstpub)
```

The number and percentage of papers that were registered beforehand:
```{r}
kable(data.frame(number = length(opendata$is_register_pred[opendata$is_register_pred == TRUE]),
           percentage = round(length(opendata$is_register_pred[opendata$is_register_pred == TRUE])/nrow(opendata)*100, 1)
           ))
```

Confidence interval for registration:
```{r}

kable(round(epi.prev(pos = length(opendata$is_register_pred[opendata$is_register_pred == TRUE]),
         tested = nrow(opendata),
         se = 0.955,
         sp = 0.997)$ap, 
      1))
```

Comparing registration rate in three years:
```{r}
kable(round(prop.table(table(opendata$is_register_pred, opendata$year_firstpub), 2)*100, 1))
```

Chi-square test:
```{r}
chisq.test(opendata$is_register_pred, opendata$year_firstpub)
```

The number and percentage of papers that shared data:
```{r}
kable(data.frame(number = length(opendata$is_open_data[opendata$is_open_data == TRUE]),
           percentage = round(length(opendata$is_open_data[opendata$is_open_data == TRUE])/nrow(opendata)*100, 1)
           ))
```

Confidence interval for data sharing:
```{r}

kable(round(epi.prev(pos = length(opendata$is_open_data[opendata$is_open_data == TRUE]),
         tested = nrow(opendata),
         se = 0.758,
         sp = 0.986)$ap, 
      1))
```

Comparing data sharing rate in three years:
```{r}
kable(round(prop.table(table(opendata$is_open_data, opendata$year_firstpub), 2)*100, 1))
```

Chi-square test:
```{r}
chisq.test(opendata$is_open_data, opendata$year_firstpub)
```

The number and percentage of papers that shared code:
```{r}
kable(data.frame(number = length(opendata$is_open_code[opendata$is_open_code == TRUE]),
           percentage = round(length(opendata$is_open_code[opendata$is_open_code == TRUE])/nrow(opendata)*100, 1)
           ))
```

Confidence interval for sharing code:
```{r}

kable(round(epi.prev(pos = length(opendata$is_open_code[opendata$is_open_code == TRUE]),
         tested = nrow(opendata),
         se = 0.587,
         sp = 0.997)$ap, 
      1))
```

Comparing sharing code rate in three years:
```{r}
kable(round(prop.table(table(opendata$is_open_code, opendata$year_firstpub), 2)*100, 1))
```

Chi-square test:
```{r}
chisq.test(opendata$is_open_code, opendata$year_firstpub)
```

Sum of five indicators:
```{r}
opendata %>% mutate(sum = rowSums(opendata %>% select(is_coi_pred, is_register_pred, is_fund_pred, is_open_data, is_open_code))) %>% select(sum) %>% tbl_summary()
```

## Journal-related differences in transparency (Table 1)

Total number of papers published in journals with a JIF:
```{r}
table(is.na(opendata$jif2020))
```

Five highest impact journals
```{r}

tapply(opendata$jif2020, opendata$journalTitle, max) %>% as.data.frame() %>% arrange(desc(.)) %>% head(5)

opendata$jif2020c = ifelse(opendata$jif2020 < 79.321,"Other", opendata$journalTitle)
```

Table 1A
```{r}
set.seed(100)
opendata %>% 
  select(journalTitle,is_coi_pred,is_fund_pred,is_register_pred,is_open_data,is_open_code) %>%
        mutate(journalTitle = fct_lump(journalTitle,n=5)) %>%
  tbl_summary(by=journalTitle, percent="column",
              label = c(is_coi_pred ~ "COI disclosure", is_fund_pred ~ "Funding disclosure", is_register_pred ~ "Protocol registration", is_open_data ~ "Data sharing", is_open_code ~ "Code sharing")) %>%
  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE)) %>%
        as_flex_table()
```

Table 1B
```{r}
set.seed(100)
opendata %>% 
  select(jif2020c,is_coi_pred,is_fund_pred,is_register_pred,is_open_data,is_open_code) %>%
  tbl_summary(by=jif2020c, percent="column",
              label = c(is_coi_pred ~ "COI disclosure", is_fund_pred ~ "Funding disclosure", is_register_pred ~ "Protocol registration", is_open_data ~ "Data sharing", is_open_code ~ "Code sharing")) %>%
  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE)) %>%
        as_flex_table()
```

## Publisher-related differences in transparency (Table 2)

Table 2
```{r}
set.seed(100)
opendata %>% 
  select(scimago_publisher,is_coi_pred,is_fund_pred,is_register_pred,is_open_data,is_open_code) %>%
        mutate(scimago_publisher = fct_lump(scimago_publisher,n=6)) %>%
  tbl_summary(by=scimago_publisher, percent="column",
              label = c(is_coi_pred ~ "COI disclosure", is_fund_pred ~ "Funding disclosure", is_register_pred ~ "Protocol registration", is_open_data ~ "Data sharing", is_open_code ~ "Code sharing")) %>%
  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE)) %>%
        as_flex_table()
```

## Journal Impact Factor (JIF) and citation count analyses (Table 3)

Total number of papers published in journals with no JIF:
```{r}
sum(is.na(opendata$jif2020))
```
Normality of JIFs and citations
```{r}
hist(opendata$jif2020)
hist(opendata$citedByCount)
ad.test(opendata$jif2020)
ad.test(opendata$citedByCount)
```

JIF and citations based on COI disclosure:
```{r}
kable(opendata %>% 
  group_by(is_coi_pred) %>% 
  summarise(Median = round(median(jif2020, na.rm = T), 1),
            IQR = round(IQR(jif2020, na.rm = T), 2)))

kable(opendata %>% 
  group_by(is_coi_pred) %>% 
  summarise(Median = round(median(citedByCount, na.rm = T), 1),
            IQR = round(IQR(citedByCount, na.rm = T), 2)))
```

Data was not normally distributed, hence we will use unpaired two-samples Wilcoxon test.
```{r}
wilcox.test(jif2020~is_coi_pred, data = opendata, exact = FALSE)
wilcox.test(citedByCount~is_coi_pred, data = opendata, exact = FALSE)
```

JIF and citations based on funding disclosure:
```{r}
kable(opendata %>% 
  group_by(is_fund_pred) %>% 
  summarise(Median = round(median(jif2020, na.rm = T), 1),
            IQR = round(IQR(jif2020, na.rm = T), 2)))

kable(opendata %>% 
  group_by(is_fund_pred) %>% 
  summarise(Median = round(median(citedByCount, na.rm = T), 1),
            IQR = round(IQR(citedByCount, na.rm = T), 2)))
```

Data was not normally distributed, hence we will use unpaired two-samples Wilcoxon test.
```{r}
wilcox.test(jif2020~is_fund_pred, data = opendata, exact = FALSE)
wilcox.test(citedByCount~is_fund_pred, data = opendata, exact = FALSE)
```

JIF and citations based on registration:
```{r}
kable(opendata %>% 
  group_by(is_register_pred) %>% 
  summarise(Median = round(median(jif2020, na.rm = T), 1),
            IQR = round(IQR(jif2020, na.rm = T), 2)))

kable(opendata %>% 
  group_by(is_register_pred) %>% 
  summarise(Median = round(median(citedByCount, na.rm = T), 1),
            IQR = round(IQR(citedByCount, na.rm = T), 2)))
```

Data was not normally distributed, hence we will use unpaired two-samples Wilcoxon test.
```{r}
wilcox.test(jif2020~is_register_pred, data = opendata, exact = FALSE)
wilcox.test(citedByCount~is_register_pred, data = opendata, exact = FALSE)
```

JIF and citations based on data sharing:
```{r}
kable(opendata %>% 
  group_by(is_open_data) %>% 
  summarise(Median = round(median(jif2020, na.rm = T), 1),
            IQR = round(IQR(jif2020, na.rm = T), 2)))

kable(opendata %>% 
  group_by(is_open_data) %>% 
  summarise(Median = round(median(citedByCount, na.rm = T), 1),
            IQR = round(IQR(citedByCount, na.rm = T), 2)))
```

Data was not normally distributed, hence we will use unpaired two-samples Wilcoxon test.
```{r}
wilcox.test(jif2020~is_open_data, data = opendata, exact = FALSE)
wilcox.test(citedByCount~is_open_data, data = opendata, exact = FALSE)
```

JIF and citations based on code sharing:
```{r}
kable(opendata %>% 
  group_by(is_open_code) %>% 
  summarise(Median = round(median(jif2020, na.rm = T), 1),
            IQR = round(IQR(jif2020, na.rm = T), 2)))

kable(opendata %>% 
  group_by(is_open_code) %>% 
  summarise(Median = round(median(citedByCount, na.rm = T), 1),
            IQR = round(IQR(citedByCount, na.rm = T), 2)))
```

Data was not normally distributed, hence we will use unpaired two-samples Wilcoxon test.
```{r}
wilcox.test(jif2020~is_open_code, data = opendata, exact = FALSE)
wilcox.test(citedByCount~is_open_code, data = opendata, exact = FALSE)
```

## Monthly trend
Figure 1:
```{r}

proportions <- opendata %>%
        summarise("COI Disclosure" = sum(is_coi_pred == TRUE),
                  "Funding disclosure" = sum(is_fund_pred == TRUE),
                  "Protocol registration" = sum(is_register_pred == TRUE),
                  "Data sharing" = sum(is_open_data == TRUE),
                  "Code sharing" = sum(is_open_code == TRUE)) %>%
        t() %>%
        as.data.frame() %>%
        rownames_to_column(var = "indicator") %>%
        mutate(percentage = round(V1/nrow(opendata)*100, 1))

p1 <- proportions %>% ggplot() +
        aes(x = reorder(indicator, V1),
            y = V1,
            fill = indicator,
            col = indicator) +
        geom_col(alpha = 0.75) +
        geom_text(aes(label = percentage), hjust = -0.1, size = 3) +
        scale_y_continuous(
                name = "\nNumber of articles (in ten thousands)",
                breaks = seq(0, 10E4, 5E4),
                labels = seq(0, 10, 5),
                expand = expansion(mult = c(0, 0.2))
                ) +
        coord_flip() +
        xlab(NULL) +
        theme(
                legend.position = "none", 
                panel.grid.major.y = element_blank()
        )
```

For the line grapht, first, we add the real publication year and month to the datasets. The real publication year/month is the year/month the paper was first appear online stored in firstPublicationDate column.
```{r}
opendata$YearMonth <- format(as.Date(opendata$firstPublicationDate), "%Y-%m")
```

```{r}

indicator_by_year <- 
        opendata %>% 
        select(YearMonth,
               is_coi_pred,
               is_fund_pred,
               is_register_pred,
               is_open_data,
               is_open_code) %>%
        gather("indicator", "value", -YearMonth) %>%
        count(YearMonth, indicator, value) %>%
        mutate(indicator = recode(indicator,
                                  is_coi_pred = "COI Disclosure",
                                  is_fund_pred = "Funding disclosure",
                                  is_register_pred = "Protocol registration",
                                  is_open_data = "Data sharing",
                                  is_open_code = "Code sharing")) %>%
        complete(indicator, value, YearMonth, fill = list(n = 0)) %>%
        group_by(YearMonth, indicator) %>% 
        mutate(p = n / sum(n)) %>%
        filter(value) %>%
        ungroup()

p2 <-
        indicator_by_year %>% 
        ggplot() +
        aes(x = YearMonth, 
            y = p,
            group = indicator,
            color = indicator) +
        geom_line(size = 0.75) +
        scale_y_continuous(limits = c(0, 1), 
                           labels = scales::percent) +
        scale_color_discrete(name = NULL) +
        scale_fill_discrete(breaks = c("COI Disclosure",
                                       "Funding disclosure",
                                       "Protocol registration",
                                       "Data sharing",
                                       "Code sharing")) +
        labs(y = "Proportion of articles (%)\n", 
             x = "\nMonth") +
        theme(panel.grid.minor = element_blank(),
              legend.position = c(0.2, 0.9),
              axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1),
              legend.key.size = unit(0.4, 'cm')
              )

figure <- ggarrange(p1, p2,
                    ncol = 2, nrow = 1, 
                    align = "hv", common.legend = F)

# tiff("Figure.tiff", width = 14, height = 7, units = "in", res = 300)
figure
#dev.off()

```

# Appendix
## Journals

Total number of papers published in journals with an SJR:
```{r}
table(is.na(opendata$sjr2020))
```

Top five journals in SJR:
```{r}
tapply(opendata$sjr2020, opendata$journalTitle, max) %>% as.data.frame() %>% arrange(desc(.)) %>% head(5)

opendata$sjr2020c = ifelse(opendata$sjr2020 < 26.304,"Other", opendata$journalTitle)
```

Table S1
```{r}
set.seed(100)
opendata %>% 
  select(sjr2020c,is_coi_pred,is_fund_pred,is_register_pred,is_open_data,is_open_code) %>%
  tbl_summary(by=sjr2020c, percent="column",
              label = c(is_coi_pred ~ "COI disclosure", is_fund_pred ~ "Funding disclosure", is_register_pred ~ "Protocol registration", is_open_data ~ "Data sharing", is_open_code ~ "Code sharing")) %>%
  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE)) %>%
        as_flex_table()
```

Top five journals in H-index:
```{r}
tapply(opendata$scimago_hindex, opendata$journalTitle, max) %>% as.data.frame() %>% arrange(desc(.)) %>% head(5)

opendata$hindexc = ifelse(opendata$scimago_hindex < 771,"Other", opendata$journalTitle)
```

Table S2
```{r}
set.seed(100)
opendata %>% 
  select(hindexc,is_coi_pred,is_fund_pred,is_register_pred,is_open_data,is_open_code) %>%
  tbl_summary(by=hindexc, percent="column",
              label = c(is_coi_pred ~ "COI disclosure", is_fund_pred ~ "Funding disclosure", is_register_pred ~ "Protocol registration", is_open_data ~ "Data sharing", is_open_code ~ "Code sharing")) %>%
  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE)) %>%
        as_flex_table()
```

## Appendix 3:
Correct p-values (considering all levels)
```{r}
opendata %>% 
  select(journalTitle,is_coi_pred) %>%
        table() %>%
        fisher.test(simulate.p.value=TRUE)
opendata %>% 
  select(journalTitle,is_fund_pred) %>%
        table() %>%
        fisher.test(simulate.p.value=TRUE)
opendata %>% 
  select(journalTitle,is_register_pred) %>%
        table() %>%
        fisher.test(simulate.p.value=TRUE)
opendata %>% 
  select(journalTitle,is_open_data) %>%
        table() %>%
        fisher.test(simulate.p.value=TRUE)
opendata %>% 
  select(journalTitle,is_open_code) %>%
        table() %>%
        fisher.test(simulate.p.value=TRUE)
```

Code sharing for all journals
```{r}
set.seed(100)
opendata %>% 
  select(journalTitle,is_open_code) %>%
  tbl_summary(by=is_open_code, percent="row", sort = list(everything() ~ "frequency"))%>%
  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE)) %>% 
        as_flex_table()
```

Data sharing for all journals
```{r}
set.seed(100)
opendata %>% 
  select(journalTitle,is_open_data) %>%
  tbl_summary(by=is_open_data, percent="row", sort = list(everything() ~ "frequency"))%>%
  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE))%>% 
        as_flex_table()
```

Protocol registration for all journals
```{r}
set.seed(100)
opendata %>% 
  select(journalTitle,is_register_pred) %>%
  tbl_summary(by=is_register_pred, percent="row", sort = list(everything() ~ "frequency"))%>%
  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE)) %>% 
        as_flex_table()
```

COI disclosure for all journals
```{r}
set.seed(100)
opendata %>% 
  select(journalTitle,is_coi_pred) %>%
  tbl_summary(by=is_coi_pred, percent="row", sort = list(everything() ~ "frequency"))%>%
  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE))%>% 
        as_flex_table()
```

Funding disclosure for all journals
```{r}
set.seed(100)
opendata %>% 
  select(journalTitle,is_fund_pred) %>%
  tbl_summary(by=is_fund_pred, percent="row", sort = list(everything() ~ "frequency"))%>%
  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE))%>% 
        as_flex_table()
```

# Publisher-related differences in transparency
Correct p-values (considering all levels)
```{r}
opendata %>% 
  select(scimago_publisher,is_coi_pred) %>%
        table() %>%
        fisher.test(simulate.p.value=TRUE)
opendata %>% 
  select(scimago_publisher,is_fund_pred) %>%
        table() %>%
        fisher.test(simulate.p.value=TRUE)
opendata %>% 
  select(scimago_publisher,is_register_pred) %>%
        table() %>%
        fisher.test(simulate.p.value=TRUE)
opendata %>% 
  select(scimago_publisher,is_open_data) %>%
        table() %>%
        fisher.test(simulate.p.value=TRUE)
opendata %>% 
  select(scimago_publisher,is_open_code) %>%
        table() %>%
        fisher.test(simulate.p.value=TRUE)
```

Data sharing for all publishers
```{r}
set.seed(100)
opendata %>% 
select(scimago_publisher,is_open_data) %>%
tbl_summary(by=is_open_data, percent="row", sort = list(everything() ~ "frequency"))%>%
 add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE))%>% 
        as_flex_table()
```

Protocol registeration in all publishers
```{r}
set.seed(100)
opendata %>% 
select(scimago_publisher,is_register_pred) %>%
tbl_summary(by=is_register_pred, percent="row", sort = list(everything() ~ "frequency"))%>%
  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE))%>% 
        as_flex_table()
```

COI disclosure in all publishers
```{r}
set.seed(100)
opendata %>% 
select(scimago_publisher,is_coi_pred) %>%
tbl_summary(by=is_coi_pred, percent="row", sort = list(everything() ~ "frequency")) %>%
  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE))%>% 
        as_flex_table()
```

Funding disclosure in all publishers
```{r}
set.seed(100)
opendata %>% 
select(scimago_publisher,is_fund_pred) %>%
tbl_summary(by=is_fund_pred, percent="row", sort = list(everything() ~ "frequency")) %>%
add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE))%>% 
        as_flex_table()
```